Evaluating Perceptual Judgements on 3D Printed Bar Charts

### 2024-03-24 jds.rmd
Material updates for revision 2

One interesting facet of @clevelandGraphical1984 is the participant recruitment methodology: "For each experiment the subjects fell into two categories: (1) a group of females, mostly housewives, without substantial technical experience; (2) a mixture of males and females with substantial technical training and working in technical jobs. 
Most of the subjects in the position-length experiment participated in the position-angle experiment; in all cases repeat subjects judged the position-angle graphs first." 
It seems likely that the authors recruited individuals within their respective departments as well as their wives.
In order to replicate this atypical participant recruitment method in the modern era, members of the UNL Statistics department and their spouses, partners, and roommates were asked to participate in our study. 
This replicates the spirit of the recruitment method without the implicit assumptions that graduate students and professors are male, heterosexual, and have unemployed spouses.

A total of `r nrow(valid.users)` participants completed the study; demographics are shown in \Cref{fig:demographics}.
The participants provided a total of `nrow(res_all)` responses.

@@

All responses that incorrectly identified the smaller bar ($n = `r nrow(res_all) - nrow(res)`$) were removed from the study before analysis; this serves as a basic attention check and leaves `r nrow(res)` valid responses. In addition, one response was removed because the participant did not enter a valid ID for a 3D printed chart.

@@

```{r midmeans-log-errors, fig.cap='Midmeans and observed values of log absolute errors for the true ratio of bars. Summary lines are computed from raw data using a loess smooth.', message=FALSE, warning=FALSE, out.width='\\textwidth'}
plot.types = c('2D', 'Rendered 3D', '3D Printed')
names(plot.types) = c('2dDigital', '3dDigital', '3dPrint')

@@

  geom_point(data = res, aes(x = ratioLabel/100, y = response, color = factor(type, levels = 1:2, labels = c("Separated", "Adjacent")))) +
  theme_bw() + ylab("Prediction") + xlab("True Proportion Difference") + 
  xlim(c(0, 1))

>>>>

One interesting facet of @clevelandGraphical1984 is the participant recruitment methodology: "For each experiment the subjects fell into two categories: (1) a group of females, mostly housewives, without substantial technical experience; (2) a mixture of males and females with substantial technical training and working in technical jobs. 
Most of the subjects in the position-length experiment participated in the position-angle experiment; in all cases repeat subjects judged the position-angle graphs first." 
It seems likely that the authors recruited individuals within their respective departments as well as their wives.
In order to replicate this atypical participant recruitment method in the modern era, members of the UNL Statistics department and their spouses, partners, and roommates were asked to participate in our study. 
This replicates the spirit of the recruitment method without the implicit assumptions that graduate students and professors are male, heterosexual, and have unemployed spouses.
A total of `r length(unique(res_all$subject))` participants completed our study; demographics are shown in \Cref{fig:demographics}.

@@

```{r, include = F}
res_sub_count <- res_all %>% 
  group_by(subject) %>% 
  count()
```

The participants provided a total of `r nrow(res_all)` responses.
While each participant was provided with 15 trials, `r sum(res_sub_count$n < 15)` participants did not complete the entire experiment.
All responses that incorrectly identified the smaller bar ($n = `r nrow(res_all) - nrow(res)`$) were removed from the study before analysis; this serves as a basic attention check and leaves `r nrow(res)` valid responses. 
In addition, one response was removed because the participant did not enter a valid ID for a 3D printed chart.

@@

```{r midmeans-log-errors, fig.cap='Midmeans and observed values of log absolute errors for the true ratio of bars. Summary lines are computed from raw data using a loess smooth.', message=FALSE, warning=FALSE, out.width='\\textwidth', fig.height=3}
plot.types = c('2D', 'Rendered 3D', '3D Printed')
names(plot.types) = c('2dDigital', '3dDigital', '3dPrint')

@@

  geom_point(data = res, aes(x = ratioLabel/100, y = response, color = factor(type, levels = 1:2, labels = c("Separated", "Adjacent")))) +
  theme_bw() + ylab("Prediction") + xlab("True Proportion Difference") + 
  xlim(c(0, 1)) + 
  theme(legend.position = 'bottom')
